\textbf{Are regime changes always bad economics? Evidence from daily financial data}

### 2019-10-15 doc/paper.tex
Added 2016 Turkish coup attempt to appendix

Markets are said to dislike political instability. We show that this is not necessarily always the case, as unexpected changes in ruler can give markets a boost when a new regime is expected to offer a more pro-business environment than the previous one.

Coups and other types of regime changes remain common, highlighting their relevance for economic and political development. Our sample consists of 5 coups, 1 assassination, 7 resignations, and 7 instances of public protect since 2000. The Arab spring is perhaps the most notable, with protests in response to oppressive regimes and low living standards spreading throughout the Middle East in late 2010. There have also been a number of failed coups such as the 2002 coup attempt against Hugo Chavez and the 2016 failed coup in Turkey. Yet despite their frequency there is little evidence on their economic consequences.

@@

While our empirical strategy does not allow us to directly test the mechanisms that drive differences in investor response from each individual event, previous theoretical literature and examination of individual cases offer some insights. Taken together, the evidence suggests that abnormal returns tend to be positive following resignations because the regime change likely replaced ``bad'' leaders whose replacements were expected to adopt better economic policies. Our examination the failed 2002 Venezuela coup shows that coups can result in large positive returns in settings where socialist or anti-business governments are replaced with pro-business regimes. Conversely, autocratic coups may increase fear of capital appropriation or the implementation of poor economic policies. Assassinations may be merely regarded as highly uncertain events, best approximating the classic ``unstable'' event where investors have few priors. Our examination of pre-event trends in abnormal returns suggests that the positive returns we observe following resignations are not driven by investors anticipating resignations, but not coups or assassinations. CARs trend downward in the days preceding resignations, but if investors anticipated a resignation that brought a more competent leader and increased stability, pre-event CARs should also trend positive.

@@

\subsection{Additional graphical depictions of individual events}

\begin{figure}[!htb]
\includegraphics{../figs/venezuela_coup_attempt_1992.pdf}
\caption{Abnormal returns surrounding the 1992 Venezuelan coup d'etat attempt}
\label{fig:AR-Ven-1992}
\end{figure}

>>>>

While conventional wisdom suggests markets dislike political instability, we show that this is not necessarily always the case. Unexpected changes in ruler virtually always increase market volatility, but the directionality is not always negative as markets can be given a boost when a new regime is expected to offer a more pro-business environment than the previous one.

Coups and other types of regime changes remain common, highlighting their relevance for economic and political development. Our sample consists of 5 coups, 1 assassination, 7 resignations, and 7 instances of public protect since 2000. The Arab spring is perhaps the most notable, with protests in response to oppressive regimes and low living standards spreading throughout the Middle East in late 2010. There have also been a number of failed coups such as the 2002 coup attempt against Hugo Chavez and the 2016 failed coup in Turkey.\footnote{The Turkish coup attempt led to negative event-day CARs of approximately -7\%. See \autoref{fig:AR-Turkey-2016} for a visual depiction.} Yet despite their frequency there is little evidence on their economic consequences.

@@

While our empirical strategy does not allow us to directly test the mechanisms that drive differences in investor response from each individual event, previous theoretical literature and examination of individual cases offer some insights. Taken together, the evidence suggests that abnormal returns tend to be positive following resignations because the regime change likely replaced ``bad'' leaders whose replacements were expected to adopt better economic policies. Our examination of pre-event trends in abnormal returns suggests that the positive returns we observe following resignations are not driven by investors anticipating resignations, but not coups or assassinations. CARs trend downward in the days preceding resignations, but if investors anticipated a resignation that brought a more competent leader and increased stability, pre-event CARs should also trend positive. Our examination the failed 2002 Venezuela coup shows that coups can result in large positive returns in settings where socialist or anti-business governments are replaced with pro-business regimes. Conversely, autocratic coups may increase fear of capital appropriation or the implementation of poor economic policies. Assassinations may be merely regarded as highly uncertain events, best approximating the classic ``unstable'' event where investors have few priors.

@@

\subsection{Graphical depictions of additional events}

\begin{figure}[!htb]
\centering
\vspace{-0.4cm}
\includegraphics[scale=0.75]{../figs/venezuela_coup_attempt_1992.pdf}
\caption{Abnormal returns surrounding the 1992 Venezuelan coup d'etat attempt}
\label{fig:AR-Ven-1992}
\end{figure}

\begin{figure}[!htb]
\centering
\vspace{-0.8cm}
\includegraphics[scale=0.75]{../figs/turkey_coup_attempt_2016.pdf}
\caption{Abnormal returns surrounding the 2016 Turkish coup d'etat attempt}
\label{fig:AR-Turkey-2016}
\end{figure}
